---
title: "Moral Machines part 2"
author: "David Steenmeijer"
date: "2022/07/06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting started
This notebook guides the reader through the R code, used in the Master's thesis 'Moral Machines in the EU: How Civilian Preferences Can Help Shape Policy'. Created in partial fulfillment for the Master's Data Science & Society at Tilburg University.

## part two
The code in this notebook is additional to the code in the orignal markdown and only shows the analyses that were conducted for the second submission. Therefore, the EDA is not included in this file.

First, the relevant packages must be imported:

```{r packages, message=FALSE, warning=FALSE, paged.print=FALSE}
# Import packages
library('dplyr') # for data manipulation
library('ggplot2') # for data visualization 
library('tidyr') # for tidying data
library('stats') # for running statistical tests and clustering
library('dendextend') # for custom visualization of dendrograms
library('colorspace') # custom visualization
library('mclust') # plotting two dendrograms in one 
library('rstatix') #for get_summary_stats function

## added in part 2:
library('caret') # for partitioning and knn
library('ggstatsplot') # for anova's
library('RSNNS') # for MLP
library('randomForest')  # for Random Forest

## Set working directory and seed
#setwd("~/Desktop/Scriptie/Thesis")
set.seed(1)
```


```{r Load data frame}
## loading data frame
Final_Frame <- read.csv('MM_EU/Final_Frame.csv', header=TRUE, sep=',')
```

```{r Remove NAs}
## remove NAs - caret does not work if there are NAs in the dataset
complete <- complete.cases(Final_Frame)
Final_Frame <- Final_Frame[complete, ]

## sanity check: complete?
which(is.na(Final_Frame)) ## should be integer(0)
```


```{r Train-Test split}
## train_test split 80:20 by country to get even distribution in country responses
partition <- createDataPartition(y = Final_Frame$UserCountry3,
                                 p = 0.8,
                                 list = FALSE)
train_frame <- Final_Frame[partition, ]
test_frame <- Final_Frame[-partition, ]

## renaming train frame to Final_Frame so code below can be executed in the same way
Final_Frame <- train_frame
```


#### Some more data manipulation
Below, the code is used to get the aggregated values of every variable in the data set, per country.
This code is identical to the original one, but conducted on the train data 

```{r EDA15: more manipulation}
Country_Intervention <- Final_Frame %>%  
  group_by(UserCountry3) %>% 
  get_summary_stats(Intervention, type = 'mean') %>% 
  rename(Intervention = mean)
Country_NumChar <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(DiffNumberOFCharacters, type = 'mean')%>% 
  rename(DiffNumberOFCharacters = mean)
Country_Man <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(Man, type = 'mean')%>% 
  rename(Man = mean)
Country_Woman <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(Woman, type = 'mean')%>% 
  rename(Woman = mean)
Country_Pregnant <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(Pregnant, type = 'mean')%>% 
  rename(Pregnant = mean)
Country_Stroller <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(Stroller, type = 'mean')%>% 
  rename(Stroller = mean)
Country_OldMan <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(OldMan, type = 'mean')%>% 
  rename(OldMan = mean)
Country_OldWoman <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(OldWoman, type = 'mean')%>% 
  rename(OldWoman = mean)
Country_Boy <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(Boy, type = 'mean')%>% 
  rename(Boy = mean)
Country_Girl <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(Girl, type = 'mean')%>% 
  rename(Girl = mean)
Country_Homeless <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(Homeless, type = 'mean')%>% 
  rename(Homeless = mean)
Country_LargeWoman <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(LargeWoman, type = 'mean')%>% 
  rename(LargeWoman = mean)
Country_LargeMan <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(LargeMan, type = 'mean')%>% 
  rename(LargeMan = mean)
Country_Criminal <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(Criminal, type = 'mean')%>% 
  rename(Criminal = mean)
Country_MaleAth <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(MaleAthlete, type = 'mean')%>% 
  rename(MaleAthlete = mean)
Country_MaleExec <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(MaleExecutive, type = 'mean')%>% 
  rename(MaleExecutive = mean)
Country_MaleDoctor <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(MaleDoctor, type = 'mean')%>% 
  rename(MaleDoctor = mean)
Country_FemaleExec <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(FemaleExecutive, type = 'mean')%>% 
  rename(FemaleExecutive = mean)
Country_FemaleAth <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(FemaleAthlete, type = 'mean')%>% 
  rename(FemaleAthlete = mean)
Country_FemaleDoctor <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(FemaleDoctor, type = 'mean')%>% 
  rename(FemaleDoctor = mean)
Country_Cat <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(Cat, type = 'mean')%>% 
  rename(Cat = mean)
CountryDog <- Final_Frame %>% 
  group_by(UserCountry3) %>% 
  get_summary_stats(Dog, type = 'mean')%>% 
  rename(Dog = mean)

Country_Averaged <- full_join(Country_Intervention, Country_NumChar, 
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE) %>% 
  select(-variable.x, -variable.y)
# add Man 
Country_Averaged <- full_join(Country_Averaged, Country_Man,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add Woman
Country_Averaged <- full_join(Country_Averaged, Country_Woman,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add Pregnant
Country_Averaged <- full_join(Country_Averaged, Country_Pregnant,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add Stroller
Country_Averaged <- full_join(Country_Averaged, Country_Stroller,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add OldMan
Country_Averaged <- full_join(Country_Averaged, Country_OldMan,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add OldWoman
Country_Averaged <- full_join(Country_Averaged, Country_OldWoman,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add Boy
Country_Averaged <- full_join(Country_Averaged, Country_Boy,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add Girl
Country_Averaged <- full_join(Country_Averaged, Country_Girl,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add Homeless
Country_Averaged <- full_join(Country_Averaged, Country_Homeless,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add LargeWoman
Country_Averaged <- full_join(Country_Averaged, Country_LargeWoman,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add LargeMan
Country_Averaged <- full_join(Country_Averaged, Country_LargeMan,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add Criminal
Country_Averaged <- full_join(Country_Averaged, Country_Criminal,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add MaleExecutive
Country_Averaged <- full_join(Country_Averaged, Country_MaleExec,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add MaleAthlete
Country_Averaged <- full_join(Country_Averaged, Country_MaleAth,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add MaleDoctor
Country_Averaged <- full_join(Country_Averaged, Country_MaleDoctor,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add FemaleExecutive
Country_Averaged <- full_join(Country_Averaged, Country_FemaleExec,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add FemaleAthlete
Country_Averaged <- full_join(Country_Averaged, Country_FemaleAth,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add FemaleDoctor
Country_Averaged <- full_join(Country_Averaged, Country_FemaleDoctor,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add Dog
Country_Averaged <- full_join(Country_Averaged, CountryDog,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)
# add Cat
Country_Averaged <- full_join(Country_Averaged, Country_Cat,
                              by = c('UserCountry3' = 'UserCountry3', 'n' = 'n'),
                              keep = FALSE)
Country_Averaged <- Country_Averaged %>% select(-variable)

#write.csv(Country_Averaged, 'MM_EU//Country_Averaged.csv')
```


## Analysis
In the following part, the analysis conducted to answer the research question is done.
This starts with clustering the data, finding the optimal number of clusters for this data.
Clustering is done with Wards method (Ward.D2 in the stats package).
To get robust outcomes, two different distances are used to check whether they influence the outcome: Euclidean and Manhattan distance.

```{r Analysis1: first cluster}
# First cluster: Ward.D2, Euclidean distance on 20 variables (intervention + all attributes, diffnumcharachters omitted)

Country_Averaged <- Country_Averaged %>% select(-DiffNumberOFCharacters)
distances_Eucl <- dist(Country_Averaged[3:23], method = 'euclidean')
Cluster_D2_Eucl <- hclust(distances_Eucl, method = 'ward.D2')
Labels_Order_D2_Eucl <- Country_Averaged$UserCountry3[Cluster_D2_Eucl$order]
scaled_data = as.matrix(scale(Country_Averaged[3:23]))
kmm = kmeans(scaled_data, 9, nstart=50, iter.max = 15)
kmm
```


#### Elbow Method
Elbow method is traditionally used to find the optimal k for the data:
This is done to prevent over- and underfitting.

```{r Analysis 2: Elbow method}
# Elbow method for k = 2 - 15
k.max <- 15
data <- scaled_data
wss <- sapply(1:k.max, 
              function(k){kmeans(data, k, nstart=50,iter.max = 10 )$tot.withinss})
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares",
     main= 'Elbow Method')
```

Based on the elbow method, K = 7 was chosen as the optimal number of clusters.

### K-means for with K=7 on Country_Averaged with Hartigan-Wong algorithm

```{r K-means}
colnames(Country_Averaged)
Country_Averaged <- Country_Averaged %>% select(-n)

set.seed(1)
kmeans_cluster <- kmeans(Country_Averaged[2:22], 7, iter.max = 25, algorithm = "Hartigan-Wong")
kmeans_cluster["cluster"]


```



#### Dendrogram 1: Euclidean distance, k = 7, Ward's method
\

```{r Analysis 3: dendrogram 1 euclidean}
dendro1 <- as.dendrogram(Cluster_D2_Eucl)
Labels_Order_D2_Eucl <- Country_Averaged$UserCountry3[Cluster_D2_Eucl$order]
labels(dendro1) <- Labels_Order_D2_Eucl
plotje <- set(dendro1, "labels_cex") %>% 
  set("labels_col", value = c(1:7), k=7) %>% 
  set("branches_lwd", 2) %>% 
  set("branches_k_color", value = 1:7, k = 7) %>% 
  plot(main = "Clustered Member States \nWard.D2, Eucl")

dendro1 %>% get_nodes_attr('height') ## get node's height for every split
```


#### Dendrogram 2: Manhattan distance, k = 7, Ward's method
\
```{r Analysis 4: dendrogram 2 manhattan }
distances_Manh <- dist(Country_Averaged[3:23], method='manhattan', labels(Country_Averaged$UserCountry3))
cluster_d2_manh <- hclust(distances_Manh, method = 'ward.D2')
dendro2 <- as.dendrogram(cluster_d2_manh)
Labels_Order_D2_manh <- Country_Averaged$UserCountry3[cluster_d2_manh$order]
labels(dendro2) <- Labels_Order_D2_manh
plotje_manh <- set(dendro2, "labels_cex") %>% 
  set("labels_col", value = c(1:7), k=7) %>% 
  set("branches_lwd", 2) %>% 
  set("branches_k_color", value = 1:7, k = 7) %>% 
  plot(main = "Clustered Member States \nWard.D2, Manh")
```

#### Comparing the two dendrograms
As can be seen in the plots, the clusters themselves are quite stable.
The only difference is that the place where they split off are different.
In other words, the between clusters are stable, the within clusters differ a bit.

```{r Analysis 5: comparing dendrograms, fig.width=7, fig.height=8}
## Compare the two dendrograms using dendextend
dendlist(dendro1, dendro2) %>% 
  untangle(method='step1side') %>%
  tanglegram( ## plots two dendrograms to visually compare them
    highlight_distinct_edges = FALSE,
    highlight_branches_lwd = FALSE)

dendlist(dendro1, dendro2) %>% 
  untangle(method='step1side') %>% 
  entanglement() ## produces alignment quality, the lower the better

## baker gamma is a correlation coefficient.
cor_bakers_gamma(dendro1, dendro2)
```

```{r}

kmeans_cluster['cluster']

```



Assign clusters to data frame
```{r Assign clusters to df}
countries <- c('AUT','BEL','BGR','CYP','CZE',
               'DEU','DNK','ESP','EST','FIN',
               'FRA','GRC','HRV','HUN','IRL',
               'ITA','LTU','LUX','LVA','MLT',
               'NLD','POL','PRT','ROU','SVK',
               'SVN','SWE')

## cluster solution hierarchical Ward k=7
cluster1 <- c(6,2,4,5,5,
             6,6,5,7,7,
             1,5,5,2,5,
             2,4,6,7,3,
             6,6,7,5,2,
             7,7)
cluster1

## assign labels to df from kmeans solution
cluster2 <- c(2,7,1,4,1,
              2,2,5,6,2,
              7,5,5,7,5,
              7,1,5,1,3,
              2,2,6,5,7,
              6,6)

assigned_clusters <- data.frame('UserCountry3' = countries, 'cluster1' =as.factor(cluster1), 'cluster2'=as.factor(cluster2))
```

test on small dataframe
```{r Test on small df}
# ## test on small dataframe, works!
# mini_test_frame <- test_frame[1:25, ]
# test_clusters <- full_join(mini_test_frame, assigned_clusters, by = 'UserCountry3')
# test_clusters <- test_clusters[1:25, ]
```


```{r}
# which(is.na(test_frame_clusters))
# write.csv(test_frame_clusters, 'test_frame_clusters.csv')

#write.csv(Final_Frame_Labels, 'Final_Frame_Labels.csv')
```

assign labels to data frame
```{r}
## assign labels to final frame
Final_Frame_Labels <- full_join(Final_Frame, assigned_clusters, by='UserCountry3')
```


Set train control for cross-validation
```{r Cross-calidation: train control}
## set train control hyperparameters
tr_ctrl <- trainControl(method = "repeatedcv",
                                 number = 7,
                                 repeats = 3,
                        verboseIter = TRUE,
                        sampling = 'down')
```


## look into the class imbalances
```{r Inspect class imbalances}
label_props_cluster1 <- Final_Train_Frame %>%
  group_by(cluster1) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

label_props_cluster1

label_props_cluster2 <- Final_Train_Frame %>%
  group_by(cluster2) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

label_props_cluster2

#create pie chart
ggplot(label_props_cluster1, aes(x="", y=freq, fill=cluster1)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  geom_text(aes(label = paste0(freq, "%")), position =  position_stack(vjust=0.5)) +
  labs(x = NULL, y = NULL) +
  theme_classic() +
  theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) +
  scale_fill_brewer(palette="Blues")

```



## SVM linear
Done on a small sample (1%) of the data to test whether it works and save computational power

```{r SVM1}

SVM_Model1 <- caret::train(Final_Frame_Labels[1:248000,7:26], 
                    as.factor(Final_Frame_Labels[1:248000,27]),
                    method = 'svmLinear',
                    preProcess = c("center", "scale"),
                    trControl = tr_ctrl,
                    tuneGrid = expand.grid(C=1))
SVM_Model1
```

```{r SVM2}
SVM_Model2 <- caret::train(Final_Frame_Labels[1:248000,7:26], 
                    as.factor(Final_Frame_Labels[1:248000,28]),
                    method = 'svmLinear',
                    preProcess = c("center", "scale"),
                    trControl = tr_ctrl,
                    tuneGrid = expand.grid(C=1))

SVM_Model2
```
## Multi-Layer Perceptron

```{r MLP1}
MLP_Model1 <- caret::train(x = Final_Frame_Labels[1:24800,7:26], 
                    y = as.factor(Final_Frame_Labels[1:24800,27]),
                    method = 'mlp',
                    trControl = tr_ctrl,
                    tuneGrid = expand.grid(size=(1))
                   )
MLP_Model1
```

```{r MLP2}
MLP_Model2 <- caret::train(x = Final_Frame_Labels[1:248000,7:26], 
                    y = as.factor(Final_Frame_Labels[1:248000, 28]),
                    method = 'mlp',
                    trControl = tr_ctrl,
                    tuneGrid = expand.grid(size=(1))
                   )
MLP_Model2
```

## random forest
```{r RF1}
rf_model1 <- caret::train(x = Final_Frame_Labels[1:27000,7:26], 
                    y = as.factor(Final_Frame_Labels[1:27000,27]),
                    method = 'rf',
                    trControl = tr_ctrl,
                    tuneGrid = expand.grid(mtry=2)
                   )

rf_model1

```

```{r RF2}
rf_model2 <- caret::train(x = Final_Frame_Labels[1:27000,7:26], 
                    y = as.factor(Final_Frame_Labels[1:27000,28]),
                    method = 'rf',
                    trControl = tr_ctrl,
                    tuneGrid = expand.grid(mtry=2)
                   )

rf_model2
```


## KNN
```{r KNN1}
knn_model1 <- caret::train(x = Final_Frame_Labels[1:248000,7:26], 
                    y = as.factor(Final_Frame_Labels[1:248000,27]),
                    method = 'knn',
                    trControl = trainControl(method = "repeatedcv",
                                 number = 50,
                                 repeats = 3),
                    tuneGrid=data.frame(k=1)
                   )

knn_model1
```

```{r KNN2}
knn_model2 <- caret::train(x = Final_Frame_Labels[1:248000,7:26], 
                    y = as.factor(Final_Frame_Labels[1:248000,28]),
                    method = 'knn',
                    trControl = trainControl(method = "repeatedcv",
                                 number = 50,
                                 repeats = 3),
                    tuneGrid=data.frame(k=1)
                   )

knn_model2

```


```{r RF on countries}
### run RF model to predict countries
rf_model_countries <- caret::train(x = Final_Train_Frame[,8:27], ## 
                          y = as.factor(Final_Train_Frame[,'UserCountry3']),
                          method = 'rf',
                          trControl = tr_ctrl,
                          tuneGrid = expand.grid(mtry=2))


rf_model_countries

# predict labels on test data
rf_fit_countries <- predict(rf_model_countries, newdata = Final_Test_Frame)

confusionMatrix(reference = as.factor(Final_Test_Frame[,'UserCountry3']), 
                data = as.factor(rf_fit_countries))
```


## prediction with RF
```{r Assign clusters to test data}
## assign clusters to test data
test_frame_clusters <- full_join(test_frame, assigned_clusters, by = 'UserCountry3')
```


```{r}
rf1_fit <- predict(rf_model1, newdata = Final_Test_Frame[,8:28]) # columns jumped 1 place up because of join

confusionMatrix(reference = as.factor(Final_Test_Frame[,28]), 
                data = as.factor(rf1_fit))

rf2_fit <- predict(rf_model2, newdata = Final_Test_Frame[,8:29])

confusionMatrix(reference = as.factor(Final_Test_Frame[,29]), 
                data = as.factor(rf2_fit))
```


Append clusters to data frame
```{r EU_data}
# EU_data <- read.csv('MM_EU/EU_data.csv', header = TRUE, sep = ',') 
# EU_info2 <- read.csv('EU_info2.csv', header = TRUE, sep = ';')
# 
# EU_data <- full_join(EU_data, EU_info2, by = "UserCountry3")
```

Create data frame with religiousness added
```{r new data frame with religiousness}
EU_data_full <- read.csv('EU_Info_full.csv', header = TRUE, sep = ';')

# arrange
EU_data_full <- EU_data_full %>% arrange(UserCountry3)

# add clusters to the EU_data
EU_data <- full_join(EU_data_full, assigned_clusters, by = "UserCountry3")
head(EU_data)


```


## Analysis: explain clusters
This was done for cluster1 and cluster2
To save room, only cluster1 is shown, to get the results for cluster2: change x = cluster1 to x = cluster2

```{r anova GDP, message=FALSE, warning=FALSE}
## Anova clusters on GDP MODEL 1
ggbetweenstats(
  data = EU_data,
  x = cluster1,
  y = GDP,
  type = 'nonparametric',
  var.equal = FALSE,
  pairwise.display = 's'
)

ggbetweenstats(
  data = EU_data,
  x = cluster2,
  y = GDP,
  type = 'nonparametric',
  var.equal = FALSE,
  pairwise.display = 's'
)
```

```{r anova EU years, message=FALSE, warning=FALSE, include=TRUE}
## Anova clusters on EU membership years
ggbetweenstats(
  data = EU_data,
  x = cluster1,
  y = EU_Years.2019.,
  type = 'nonparametric',
  var.equal = FALSE
)

ggbetweenstats(
  data = EU_data,
  x = cluster2,
  y = EU_Years.2019.,
  type = 'nonparametric',
  var.equal = FALSE
)
```

```{r anova gender pay gap, include=TRUE}
## Anova clusters on Gender Pay Gap
ggbetweenstats(
  data = EU_data,
  x = cluster1,
  y = Gender_Pay_Gap,
  type = 'nonparametric',
  var.equal = FALSE
)

ggbetweenstats(
  data = EU_data,
  x = cluster2,
  y = Gender_Pay_Gap,
  type = 'nonparametric',
  var.equal = FALSE
)
```

```{r anova High Tech Employment, include=TRUE}
## Anova clusters on High Tech Employment
ggbetweenstats(
  data = EU_data,
  x = cluster1,
  y = High_Tech_Employment,
  type = 'nonparametric',
  var.equal = FALSE
)

ggbetweenstats(
  data = EU_data,
  x = cluster2,
  y = High_Tech_Employment,
  type = 'nonparametric',
  var.equal = FALSE
)
```

```{r anova Population}
## Anova clusters on Population
ggbetweenstats(
  data = EU_data,
  x = cluster1,
  y = Population_2019,
  type = 'nonparametric',
  var.equal = FALSE
)

ggbetweenstats(
  data = EU_data,
  x = cluster2,
  y = Population_2019,
  type = 'nonparametric',
  var.equal = FALSE
)
```


```{r anova Population, Gini, Life_Expectancy}


## ANOVA Gini2019
ggbetweenstats(
  data = EU_data,
  x = cluster1,
  y = Gini_2019,
  type = 'nonparametric',
  var.equal = FALSE
)

## ANOVA Gini2019
ggbetweenstats(
  data = EU_data,
  x = cluster2,
  y = Gini_2019,
  type = 'nonparametric',
  var.equal = FALSE
)


```

```{r}
## ANOVA Life_Expectancy
ggbetweenstats(
  data = EU_data,
  x = cluster1,
  y = Life_Expectancy,
  type = 'nonparametric',
  var.equal = FALSE
)

ggbetweenstats(
  data = EU_data,
  x = cluster2,
  y = Life_Expectancy,
  type = 'nonparametric',
  var.equal = FALSE
)
```


```{r}
## ANOVA on religiousness
ggbetweenstats(
  data = EU_data,
  x = cluster1,
  y = religiousness,
  type = 'nonparametric',
  var.equal = FALSE
)

ggbetweenstats(
  data = EU_data,
  x = cluster2,
  y = religiousness,
  type = 'nonparametric',
  var.equal = FALSE
)
```

create new dataframes with only the 3 biggest clusters per model
```{r}
## Model 1
big_clusters1 <- EU_data %>% 
  filter(cluster2 == 2 |
           cluster2 == 5 |
           cluster2 == 7)
```


```{r anova big clusters}
## anova with 3 biggest clusters on GDP
ggbetweenstats(
  data = big_clusters1,
  x = cluster2,
  y = GDP,
  type = 'nonparametric',
  var.equal = FALSE,
  pairwise.display = 'all'
)

## EU Years
ggbetweenstats(
  data = big_clusters1,
  x = cluster2,
  y = EU_Years.2019.,
  type = 'nonparametric',
  var.equal = FALSE,
  pairwise.display = 'all'
)

## High_Tech_Employment
ggbetweenstats(
  data = big_clusters1,
  x = cluster2,
  y = High_Tech_Employment,
  type = 'nonparametric',
  var.equal = FALSE,
  pairwise.display = 'all'
)

## Gender Pay Gap
ggbetweenstats(
  data = big_clusters1,
  x = cluster2,
  y = Gender_Pay_Gap,
  type = 'nonparametric',
  var.equal = FALSE,
  pairwise.display = 'all'
)
## Population
ggbetweenstats(
  data = big_clusters1,
  x = cluster2,
  y = Population_2019,
  type = 'nonparametric',
  var.equal = FALSE,
  pairwise.display = 'all'
)

## Gini2019
ggbetweenstats(
  data = big_clusters1,
  x = cluster2,
  y = Gini_2019,
  type = 'nonparametric',
  var.equal = FALSE
)

## Life_Expectancy
ggbetweenstats(
  data = big_clusters1,
  x = cluster2,
  y = Life_Expectancy,
  type = 'nonparametric',
  var.equal = FALSE
)

## Religiousness
ggbetweenstats(
  data = big_clusters1,
  x = cluster2,
  y = religiousness,
  type = 'nonparametric',
  var.equal = FALSE
)
```

